### 2022-03-04 paper/localpsf.tex
editing writing

\subsection{Overview of the method}
The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
\begin{equation*}
\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
\end{equation*}
denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
\begin{equation*}
\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
\end{equation*}
denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
\begin{equation*}
\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
\end{equation*}
denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively. In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).

@@

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in Section SECTION, but the key principle behind the interpolation is an assumption that we call \emph{local mean-displacement invariance}. Specifically, we make the assumption that
\begin{equation}
\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
\end{equation}
if $x_i$ is close to $x$. This is shown in Figure FIG, and discussed in greater detail in Section SEC. We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the a small number of sample points $x_i$ near $x$. The weights in the linear combination are determined via radial basis function interpolation.

@@

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}

@@

		Form convolution kernels, $\convkernel_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})

		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})

		Compute all weighting functions $w_i$ using Algorithm

		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})

		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})

@@

We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs. The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
\begin{equation*}
	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
\end{equation*}
We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
\begin{equation}
	\label{eq:dirac_comb_H_action}
	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
\end{equation}
The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
%\begin{equation}
%\label{eq:phi_b}
%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
%\begin{equation}
%	\label{eq:phi_b}
%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\impulseresponse_{x_i}(z) / V(x_i) = 
	\begin{cases}
		\combresponse_k(z), & z \in E_{x_i} \\
		0, & \text{otherwise}
	\end{cases}
\end{equation}
for all $x_i \in \pointbatch_k$. By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.

@@

We now define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
\begin{equation}
	\label{eq:defn_of_Akerpcmesh}
	\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.

@@

\subsection{Hierarchical matrix construction}

>>>>

\subsection{Overview of the method}
The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
\begin{equation*}
\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
\end{equation*}
denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
\begin{equation}
\label{eq:scaled_impulse_response}
\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
\end{equation}
denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
\begin{equation*}
\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
\end{equation*}
denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively. In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).

@@

Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation. The details of the interpolation will be described in Section SECTION, but the key principle behind the interpolation is an assumption that we call \emph{local mean-displacement invariance}. Specifically, we make the assumption that
\begin{equation}
\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
\end{equation}
if $x_i$ is close to $x$. This is shown in Figure FIG. We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the a small number of sample points $x_i$ near $x$. The weights in the linear combination are determined via radial basis function interpolation.

@@

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
		\Output{H-matrix $\mathbf{A}_H$ (optionally, $\mathbf{A}_H^{\text{sym}+}$)}

@@

		Form H-matrix approximation $\mathbf{\Phi}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})

		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section \ref{sec:h_matrix_construction_short})

		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Section \ref{sec:make_spd})

@@

We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs. The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
\begin{equation*}
	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
\end{equation*}
We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
\begin{equation}
	\label{eq:dirac_comb_H_action}
	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
	=\sum_{x_i \in \pointbatch_k} \widehat{\impulseresponse}_{x_i}.
\end{equation}
The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\widehat{\impulseresponse}_{x_i}$ in \eqref{eq:scaled_impulse_response} and \eqref{eq:impulse_response_delta_action}.
%\begin{equation}
%\label{eq:phi_b}
%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
%\begin{equation}
%	\label{eq:phi_b}
%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
%\end{equation}
Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\impulseresponse_{x_i}(z) / V(x_i) = \widehat{\impulseresponse}_{x_i}(z) =
	\begin{cases}
		\combresponse_k(z), & z \in E_{x_i} \\
		0, & \text{otherwise}
	\end{cases}
\end{equation}
for all $x_i \in \pointbatch_k$. By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.

@@

We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
\begin{equation}
	\label{eq:defn_of_Akerpcmesh}
	\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
\end{equation}
The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.

@@

\subsection{Hierarchical matrix construction}
\label{sec:h_matrix_construction_short}
